---
title: "Data Preparation 2"
author: "Aje Joshua Ayodele"
date: "6 March 2020"
output: html_document
---


```{r packages}
#load libraries
library(haven) #import stata files
library(tidyverse) #for data preprocessing and visualization
library(questionr)
library(stargazer) # for reporting model summary
library(ggplot2) # for awesome visualizations
library(dagitty) # package to draw DAGs
library(ggdag) # another package to draw DAGS
library(estimatr) # summarize data sets
```

```{r functions}
# Make list of variables
  makeVlist <- function(dta) { 
  labels <- sapply(dta, function(x) attr(x, "label"))
  tibble(name = names(labels),
         label = labels)
  }
```



## Data Import
```{r data import, warning=FALSE, message=FALSE}
#import persoanl pfad data
pfad <- read_dta("ppfad.dta")
pfadLabels <- makeVlist(pfad)

#import household path dataset
hpfad <- read_dta("hpfad.dta") 
hpfadLabels <- makeVlist(hpfad) 

#import generated dataset for personal dataset
personal_gen <- read_dta("bhpgen.dta")

#import generated dataset for bhh
hh_gen <- read_dta("bhhgen.dta")
#hh_genList <- makeVlist(hh_gen)

# import bhkind (child) dataset
child <- read_dta("bhkind.dta") 
#kidLabels <- makeVlist(kid)

#import bioagel (parent-child) dataset
bioage <- read_dta("bioagel.dta") 

#import bhpequiv
bhp_equiv <- read_dta("bhpequiv.dta")
equivLabels <- makeVlist(bhp_equiv)
```

### Quick experimentation to get the age of the youngest child in a household
```{r}
# get required variables from pfad
pfad2 <- pfad %>% select(persnr, bhhhnr, gebjahr) %>% 
  mutate(age = 2017 - gebjahr) %>% arrange(persnr)

# get required variables from bioage
bioage2 <- bioage %>%  select(persnr, hhnrakt)

#merge pfad and bioage
bioage2 <- merge(pfad2, bioage2, by = ("persnr")) %>% 
  group_by(hhnrakt) %>% 
  mutate(age_youngest = min(age)) %>%
  ungroup()  %>% 
  select(bhhhnr, age_youngest) %>% 
  filter(bhhhnr != -2) %>% 
  distinct(bhhhnr, .keep_all = TRUE) %>% 
  arrange(bhhhnr)

# get required variables from kid
child2 <- child %>% select(persnr, hid, bhkgjahr, bhkzahl)

# calculate age of youngetst child per household
child2 <- merge(pfad2, child2, by = ("persnr")) %>% 
  select(bhhhnr, age, bhkzahl) %>% 
  group_by(bhhhnr) %>% 
  mutate(age_youngest = min(age)) %>% 
  ungroup() %>% 
  distinct(bhhhnr, .keep_all = TRUE) %>% 
  select(bhhhnr, age_youngest, bhkzahl) %>% 
  arrange(bhhhnr)

# child2 contains households not in bioage2 and vice-versa, merge both

child3 <- full_join(child2, bioage2, by = "bhhhnr") %>% arrange(bhhhnr)

# fill NA values in age_youngest.x with corresponding values in age_youngest.y
for (i in 1:nrow(child3)) {
  if (is.na(child3[i, "age_youngest.x"])) {
    child3[i, "age_youngest.x"] <- child3[i, "age_youngest.y"]
  }
}

child3 <- child3 %>% select(-age_youngest.y) %>% 
  rename(age_youngest = "age_youngest.x")

#save dataset
write.csv(child3, file = "child3.csv", row.names = FALSE)

#remove non-required datasets
rm(pfad2, bioage2, child2)
```

## Variable Selection
We select required variables from the respective datasets
```{r}
# PFAD
core_p <- pfad %>% 
  select(persnr, sex, gebjahr, psample, bhhhnr, bhnetto, bhsampreg, bhpop,
        immiyear, germborn, corigin, migback) %>% 
  filter(!psample %in% c(17, 18, 19), #exclude refugees
         between(bhnetto, 10, 19), #successful interviews
         bhpop %in% c(1,2)) %>% #private households (1- German HH-head; 2: Foreign)
  mutate(age = 2017 - gebjahr) %>% 
  filter(between(age, 25, 64)) %>% 
  arrange(persnr)


# GENERATED PERSONAL DATASET
personal_gen <- personal_gen %>% 
  select(persnr, bhhhnr, bhbilzeit, bhfamstd, emplst17, 
         isced11_17, partnr17) %>% 
  rename(
    "amt_edu_yrs" = bhbilzeit,
    "marital_status" = bhfamstd,
    "emp_status" = emplst17,
    "edu_level" = isced11_17,
    "partner_persnr" = partnr17
  ) %>% 
  filter(marital_status %in% c(1, 7)) %>% #married or registered partnership living together
  arrange(persnr)

## HOUSEHOLD GENERATED DATASET
hh_gen <- hh_gen %>% select(bhhhnr, nuts117, typ1hh17, typ2hh17, hinc17) %>% 
  rename(
    state = "nuts117",
    hhtype1 = "typ1hh17",
    hhtype2 = "typ2hh17",
  ) %>% 
  arrange(bhhhnr)


# BHP_EQUIV DATASET
bhp_equiv <- bhp_equiv %>% 
  select(bhhhnr, persnr, d1110517, d1110617, d1110717,i1110117, i1110217, w1110217) %>% 
  rename(
    "rel_to_head" = d1110517,
    "hh_size" = d1110617,
    "num_kids" = d1110717,
    "pregov_income" = i1110117,
    "postgov_income" = i1110217,
    "hh_weight" = w1110217
  ) %>% 
  arrange(persnr)
```

## Data Merging

```{r}
core_p <- core_p %>% arrange(bhhhnr)

#merge the core (pfad) data and the hh_gen data
merge1 <- merge(core_p, hh_gen, by = c("bhhhnr")) %>% arrange(persnr)

# merge the newly merged data with the personal_gen dataset
merge2 <- merge(merge1, personal_gen, by = c("persnr", "bhhhnr")) %>% arrange(persnr)
rm(merge1)

#merge with bhp_equiv
merge3 <- merge(merge2, bhp_equiv, by = c("persnr", "bhhhnr")) %>% arrange(bhhhnr)
rm(merge2)
```


## Data Wrangling
```{r}
# some data wrangling
merge3 <- merge3 %>%  
  mutate(adult = hh_size - num_kids,
  other_adult = adult - 1,
  pregov_income_eq = pregov_income / (1 + other_adult * 0.5 + num_kids * 0.3),
  postgov_income_eq = postgov_income / (1 + other_adult * 0.5 + num_kids * 0.3),
  poverty_level_pre = ifelse(pregov_income_eq >= (median(pregov_income_eq) * 0.60), "Not poor", "Relatively poor"),
  poverty_level_post = ifelse(postgov_income_eq >= (median(postgov_income_eq) * 0.60), "Not poor", "Relatively poor"),
  poverty_level_pre = factor(poverty_level_pre), 
  poverty_level_post = factor(poverty_level_post),
  #num_kids = case_when(
    #num_kids == 0 ~ "No child",
    #num_kids == 1 ~ "1 child",
    #num_kids == 2 ~ "2 children",
    #TRUE ~ "3 or more children"),
  #num_kids = factor(num_kids),
  rel_to_head = case_when(
    rel_to_head == 1 ~ "Head",
    rel_to_head == 2 ~ "Partner",
    rel_to_head == 3 ~ "Child",
    rel_to_head == 4 ~ "Relative",
    TRUE ~ "Nonrelative"),
  sex = ifelse(sex == 1, "Male", "Female"),
  sex = factor(sex),
  emp_status = case_when(
  emp_status == 1 ~ "Full-time employed",
  emp_status == 2 | emp_status == 4 ~ "Part-time employed",
  TRUE ~ "Not employed"),
  emp_status = factor(emp_status),
  germborn = case_when(
    germborn == 1 & migback == 1 ~ "native",
    germborn == 1 & migback == 3 ~ "second_gen migrant",
    TRUE ~ "first_gen migrant"),
  germborn = factor(germborn),
  edu_level = case_when(
    between(edu_level, 0, 2) ~ "Low",
    edu_level %in% c(3,4) ~ "Middle",
    between(edu_level, 5, 8) ~ "High",
    TRUE ~ "NA"),
  edu_level = factor(edu_level),
  bhpop = ifelse(bhpop == 1, "German HH-Head", "Foreign HH-Head"),
        bhpop = factor(bhpop)) %>% 
  filter(!hhtype1 %in% c(3, 7, 8), # exclude other non-couple households
         !rel_to_head %in% c("Child", "Relative", "Nonrelative"), # restrict sample to household head and partner
         !partner_persnr < 0) %>% #remove persons whose spouse are not included in the survey data
  arrange(bhhhnr) 


#rearrange column order
my_df <- merge3 %>%  select(bhhhnr, persnr, sex, bhpop, rel_to_head, age, germborn, migback, corigin, edu_level, emp_status, hh_size, num_kids, partner_persnr, hhtype2, adult, other_adult, pregov_income_eq, postgov_income_eq, poverty_level_pre, poverty_level_post, hh_weight)

#get unique personal id for all individuals who both completed the survey with their spouse
partner_persnr <- my_df$partner_persnr

unique_pernsr <- sort(union(my_df$persnr, partner_persnr))

#get unique household numbers
unique_bhhhnr <- sort(unique(my_df$bhhhnr))
```


We repeat the same variable selection step having obtained the unique personal ids of my_df. The purpose of repeating this step is to use the unique personal ids to filter our pfad dataset such that the dataset contains only the individuals whose  spuse also successfuly completed the survey. 

```{r}
# PFAD: extract couples using unique_persnr
core_p_rep <- pfad %>% 
  select(persnr, sex, gebjahr, psample, bhhhnr, bhnetto, bhsampreg, bhpop,
        immiyear, germborn, corigin, migback) %>% 
  filter(persnr %in% unique_pernsr) %>%  #extract the couples using unique_persnr
  arrange(persnr)

#read in original personal generated dataset 
personal_gen_rep <- read_dta("bhpgen.dta")

# GENERATED PERSONAL DATASET: extract couples using unique_persnr
personal_gen_rep <- personal_gen_rep %>% 
  select(persnr, bhhhnr, bhbilzeit, bhfamstd, emplst17, 
         isced11_17, partnr17) %>% 
  filter(persnr %in% unique_pernsr) %>% # get the couples using unique_pernsr
  rename(
    "amt_edu_yrs" = bhbilzeit,
    "marital_status" = bhfamstd,
    "emp_status" = emplst17,
    "edu_level" = isced11_17,
    "partner_persnr" = partnr17
  ) %>% 
  arrange(persnr)

#read in original personal generated dataset 
bhp_equiv_rep <- read_dta("bhpequiv.dta")

# BHP_EQUIV DATASET: extract couples using unique_persnr
bhp_equiv_rep <- bhp_equiv_rep %>% 
  select(bhhhnr, persnr, d1110517, d1110617, d1110717,i1110117, i1110217, w1110217) %>% 
  filter(persnr %in% unique_pernsr) %>% # get the couples using unique_pernsr
  rename(
    "rel_to_head" = d1110517,
    "hh_size" = d1110617,
    "num_kids" = d1110717,
    "pregov_income" = i1110117,
    "postgov_income" = i1110217,
    "hh_weight" = w1110217
  ) %>% 
  arrange(persnr)
```


## REMERGING OF DATASETS
```{r}
core_p_rep <- core_p_rep %>% arrange(bhhhnr)

#merge the core_p_rep data and the hh_gen data
merge1 <- merge(core_p_rep, hh_gen, by = c("bhhhnr"))

merge1 <- merge1 %>% arrange(persnr)

# merge the newly merged data with the personal_gen_rep dataset
merge2 <- merge(merge1, personal_gen_rep, by = c("persnr", "bhhhnr"))
rm(merge1)
merge2 <- merge2 %>% arrange(persnr)

#merge with bhp_equiv
merge3 <- merge(merge2, bhp_equiv_rep, by = c("persnr", "bhhhnr"))
rm(merge2)
merge3 <- merge3 %>% arrange(bhhhnr)

# some data wrangling
merge3 <- merge3 %>%  
  mutate(adult = hh_size - num_kids,
  other_adult = adult - 1,
  pregov_income_eq = pregov_income / (1 + other_adult * 0.5 + num_kids * 0.3),
  postgov_income_eq = postgov_income / (1 + other_adult * 0.5 + num_kids * 0.3),
  poverty_level_pre = ifelse(pregov_income_eq >= (median(pregov_income_eq) * 0.60), "Not poor", "Relatively poor"),
  poverty_level_post = ifelse(postgov_income_eq >= (median(postgov_income_eq) * 0.60), "Not poor", "Relatively poor"),
  poverty_level_pre = factor(poverty_level_pre), 
  poverty_level_post = factor(poverty_level_post),
  rel_to_head = case_when(
    rel_to_head == 1 ~ "Head",
    rel_to_head == 2 ~ "Partner",
    rel_to_head == 3 ~ "Child",
    rel_to_head == 4 ~ "Relative",
    TRUE ~ "Nonrelative"),
  sex = ifelse(sex == 1, "Male", "Female"),
  sex = factor(sex),
  emp_status = case_when(
  emp_status == 1 ~ "Full-time employed",
  emp_status == 2 | emp_status == 4 ~ "Part-time employed",
  TRUE ~ "Not employed"),
  emp_status = factor(emp_status),
  germborn = case_when(
    germborn == 1 & migback == 1 ~ "native",
    germborn == 1 & migback == 3 ~ "second_gen migrant",
    TRUE ~ "first_gen migrant"),
  germborn = factor(germborn),
  edu_level = case_when(
    between(edu_level, 0, 2) ~ "Low",
    edu_level %in% c(3,4) ~ "Middle",
    between(edu_level, 5, 8) ~ "High",
    TRUE ~ "NA"),
  edu_level = factor(edu_level),
  bhpop = ifelse(bhpop == 1, "German HH-Head", "Foreign HH-Head"),
  bhpop = factor(bhpop),
  age = 2017 - gebjahr) %>% 
  arrange(bhhhnr) 


#rearrange column order
my_df <- merge3 %>%  select(bhhhnr, persnr, sex, bhpop, rel_to_head, age, germborn, migback, corigin, edu_level, emp_status, hh_size, num_kids, partner_persnr, hhtype2, adult, other_adult, pregov_income_eq, postgov_income_eq, poverty_level_pre, poverty_level_post, hh_weight)

#filter for couples between the age of 25 and 65
remove_age <- my_df %>% filter(!between(age, 25, 64)) %>% select(bhhhnr, persnr, age)
remove_bhhnr <- remove_age$bhhhnr

#remove household with either couple being under-25 or over-64
my_df<- my_df %>% filter(!bhhhnr %in% remove_bhhnr)

#calculate freq distribution of the unique household ids
bhhhnr_freq<- table(my_df$bhhhnr)

#retain households with duplicate unique ids (i.e. for husband and wive)
my_df <- subset(my_df, bhhhnr %in% names(bhhhnr_freq[bhhhnr_freq > 1]))

# remove same-sex couple
#get unique HH id  for the men in the sample
unq_hus <- unique(my_df[my_df$sex == "Male", "bhhhnr"])

#get unique HH id  for the women in the sample
unq_wife <- unique(my_df[my_df$sex == "Female", "bhhhnr"])

#get unique household ids
unq <- unique(my_df$bhhhnr)

# extract households where both spouses are women
same_sex1 <- setdiff(unq, unq_hus)

# extract households where both spouses are men
same_sex2 <- setdiff(unq, unq_wife)

#remove same-sex couple households
my_df <- my_df %>% filter(!bhhhnr %in% same_sex1,
                          !bhhhnr %in% same_sex2)
```



```{r}
#getting characteristics of the male spouse
husband_educ <- my_df %>% filter(sex == "Male") %>% select(bhhhnr, edu_level) %>% rename("mspouse_edu" = edu_level)
husband_age <- my_df %>% filter(sex == "Male") %>% select(bhhhnr, age) %>% rename("mspouse_age" = age)
husband_emp <- my_df %>% filter(sex == "Male") %>% select(bhhhnr, emp_status) %>% rename("mspouse_emp" = emp_status)
husband_bg <- my_df %>% filter(sex == "Male") %>% select(bhhhnr, germborn) %>% rename("mspouse_bg" = germborn)

#getting characteristics of the female spouse
wife_educ <- my_df %>% filter(sex == "Female") %>% select(bhhhnr, edu_level) %>% rename("woman_edu" = edu_level)
wife_age <- my_df %>% filter(sex == "Female") %>% select(bhhhnr, age) %>% rename("woman_age" = age)
wife_emp <- my_df %>% filter(sex == "Female") %>% select(bhhhnr, emp_status) %>% rename("woman_emp" = emp_status)
wife_bg <- my_df %>% filter(sex == "Female") %>% select(bhhhnr, germborn) %>% rename("woman_bg" = germborn)

#add this infos to the data set
my_df <- merge(my_df, husband_educ, by = "bhhhnr")
my_df <- merge(my_df, husband_age, by = "bhhhnr")
my_df <- merge(my_df, husband_emp, by = "bhhhnr")
my_df <- merge(my_df, husband_bg, by = "bhhhnr")
my_df <- merge(my_df, wife_educ, by = "bhhhnr")
my_df <- merge(my_df, wife_age, by = "bhhhnr")
my_df <- merge(my_df, wife_emp, by = "bhhhnr")
my_df <- merge(my_df, wife_bg, by = "bhhhnr")


#label the households
label_table <- my_df %>% 
  mutate(hh_label = case_when(
        mspouse_bg == "native" & woman_bg == "native" ~ "native",
        (mspouse_bg == "native" & woman_bg == "first_gen migrant") | (mspouse_bg == "first_gen migrant" & woman_bg == "native") ~ "mixed",
        (mspouse_bg == "native" & woman_bg == "second_gen migrant") | (mspouse_bg == "second_gen migrant" & woman_bg == "native") ~ "mixed",
        TRUE ~ "migrant")) %>% 
  filter(rel_to_head == "Head") %>% 
  select(bhhhnr, hh_label)


#merge label table with my_df
my_df <- merge(my_df, label_table, by = "bhhhnr")

# merge my_df with child table 

my_df2 <- left_join(my_df, child3, by = "bhhhnr")

#get final data set
final_df <- my_df2 %>% 
  distinct(bhhhnr, .keep_all = TRUE) %>%
  select(bhhhnr, mspouse_age, mspouse_edu, mspouse_emp, woman_age, woman_edu, woman_emp,
         hh_size, num_kids, age_youngest, adult, other_adult, pregov_income_eq, postgov_income_eq, poverty_level_pre, poverty_level_post, hh_label, hh_weight)

final_df <- final_df %>% 
  filter(woman_edu != "NA",
         mspouse_edu != "NA") %>% 
  mutate(
  hh_label = factor(hh_label),
  num_kids = case_when(
    num_kids == 0 ~ "childless",
    num_kids == 1 ~ "1 child",
    num_kids == 2 ~ "2 children",
    TRUE ~ "3 or more children"
  ),
  num_kids = factor(num_kids),
  pre_school = case_when(
    age_youngest >= 0 & age_youngest <=6 ~ "With",
    TRUE ~ "Without"
  ),
  pre_school = factor(pre_school),
  mspouse_age = case_when(
    mspouse_age >= 25 & mspouse_age <= 35 ~ "25-35",
    mspouse_age >= 36 & mspouse_age <= 50 ~ "36-50",
    TRUE ~ "51-64"
  ),
  mspouse_age = factor(mspouse_age),
  woman_age = case_when(
    woman_age >= 25 & woman_age <= 35 ~ "25-35",
    woman_age >= 36 & woman_age <= 50 ~ "36-50",
    TRUE ~ "51-64"
  ),
  woman_age = factor(woman_age),
  syear = 2017
) %>% 
  select(-c("age_youngest")) %>% 
  droplevels.data.frame()


#change reference categories
final_df$mspouse_edu <- fct_relevel(final_df$mspouse_edu, "Low", after = 2)
final_df$mspouse_emp <- fct_relevel(final_df$mspouse_emp, "Not employed", after = 2)
final_df$woman_edu <- fct_relevel(final_df$woman_edu, "Low", after = 2)
final_df$woman_emp <- fct_relevel(final_df$woman_emp, "Not employed", after = 2)
final_df$num_kids <- fct_relevel(final_df$num_kids, "childless")
final_df$pre_school <- fct_relevel(final_df$pre_school, "Without")

saveRDS(final_df, "final_df17.RData")
```


